---
title: "Sales Trends and Forecast"
author: "Sajjad"
date: "5/14/2022"
params:
  justContext: FALSE
  site: "ALL"
  season: "ALL"
  newstuff: TRUE
  statusReport: FALSE
  incomplete: FALSE
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,comment="##",fig.width=15, fig.height=6.5,dpi=2*72, warning=F)
options(dplyr.summarise.inform = FALSE)
```

```{r SetOutputDirectory}
# This is necessary for knit_child to work properly when using "Run" in RStudio
knitr::opts_knit$set(output.dir = ".")
options(knitr.duplicate.label = 'allow')
```

```{r, include=FALSE}
output <- knitr::knit_child(paste0(here::here(), "/Analysis/prepareData.Rmd"))
```

In this document I look at the sales trends over time.  
```{r}
library(fpp2)
```


# Sales Trends
```{r}
trends <- train %>%
  select(fullVisitorId, date, month, weekDay, transactionRevenue, year, 
         visitTimeOfDay)
```

# Time series analysis
```{r}
ts <- train %>%
  mutate(transactionRevenue = if_else(is.na(transactionRevenue), 0,
                                      as.numeric(transactionRevenue))) %>%
  group_by(date) %>%
  summarise(sales = sum(transactionRevenue)) %>%
  mutate(sales = if_else(sales == 0, 0, log(sales)))
```

```{r}
ts <- ts(ts$sales, start = c(2016, yday("2016-08-01")), 
         end=c(2017, yday("2017-08-01	")), frequency = 365)
```

```{r}
ggtsdisplay(ts, plot.type = "histogram",smooth = T)

```

We can see from the plot that there's no sales on day "2016-08-03", which is 
unusual so let's just impute this value with the mean days before and after it.

```{r}
ts <- train %>%
  mutate(transactionRevenue = if_else(is.na(transactionRevenue), 0,
                                      as.numeric(transactionRevenue))) %>%
  group_by(date) %>%
  summarise(sales = sum(transactionRevenue)) %>%
  mutate(sales = if_else(sales == 0, 0, log(sales)))

ts$sales[ts$date == "2016-08-03"] <- mean(ts$sales[ts$date == "2016-08-02"],
                                           ts$sales[ts$date == "2016-08-04"])
```

```{r}
ts <- ts(ts$sales, start = c(2016, yday("2016-08-01")), 
         end=c(2017, yday("2017-08-01")), frequency = 365)
```

```{r}
ggtsdisplay(ts, plot.type = "partial",smooth = T)
```

Now the ts looks much better. HEre's the insights from it:

+ Overal the series looks steady around the mean.
+ We have huge lag 1, 2, and 3. Every day's sales is positively correlated with the day before
same day from last week and is negatively correlated with 2 days ago.
+ We can observe a huge weekly pattern.
+ There are few drops in the sales that needs further investigation: mid 2016, beginning of 2017, 2017.


## ARIMA Forecast
```{r arima}
model <- auto.arima(ts)
forecast <- forecast(model, h=30)
```

```{r}
autoplot(forecast)
```